---
title: " 'HS Team Data' Planning and Resource Document"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Objectives

*  Collaboratively develop "JS Team Data", a Data Science Team at John Sweet High School in Crockett, California
*  Introduce interested students to the "hot" field of Data Science
*  Learn together how to make web sites and other data-related resources that can have an impact and/or are just fun
*  Team "meetings" might include
    *  Presentation of a new tool or topic 
    *  Discussion of the topic
    *  Working together on a current project

### Activities

*  The Team will decide together what tools (e.g. software) to use moving forward, and decide what web applications and/or other "products" to build based on interests and motivations of team members
* Core team focus will be on retrieving/collecting data, organizing and curating data, and analyzing and displaying data, BUT there would be some great team roles for other students/members like artists (design of pictures and logos), school leaders (suggesting directions for team projects), and others.
*  Eventually, the team could apply for some small consulting gigs, to learn more about coding in the "real world" and to build some team resources

### Mentor or "Coach"
*  Team Mentor (Coach?), Michael Samuel, will introduce the Team to a suite of tools based on the free and open source R ecosystem, including
     *  R base software,
     *  RStudio (graphical user interface)
     *  tidyverse set of packages for "data wrangling"
     *  R Shiny package for interactive visual display
     *  tmap for mapping
     *  and many others
*  Michael will also introduce the use of GitHub as the repository for Team work, and to facilitate several other aspects of team collaboration
* A bit more about Michael can be found [here](http://dlab.berkeley.edu/people/dr-michael-samuel) or [here](https://www.linkedin.com/in/mcsamuel/)

### Logistics

* Some logistics will need to be determined based on the schedules of interested students
* A tentative plan would be to meet one afternoon a week after school in the ??? room, and one afternoon a week "virtually" (i.e. using WebEx), with the mentor at home or at work, and the students at school or other convenient locations
* With the students we may well change the name from "Team" to something more exciting.  The idea of the "Team" is much less about competition with other schools (although maybe that's an idea too...) but about working as a team together to learn, have fun, and maybe make some cool and/or useful things together
   

### GitHub Site
* Initial Team GitHub Site is [here](https://github.com/mcSamuelDataSci/Data_Science_Team)
*  The document you are reading, in both HTML and .docx formats, can be found on this site, as well at the "R Markdown" file that created these documents (DS_team_introduction.Rmd)
*  All other current Team files can also be found here


### "Open Data and Community Health"


* We will emphasize using "Open Data" and other data that is freely available in a machine readable format.  The mentor will introduce these types of data focusing on community health and "Social Determinants of Health", but after that the Team could shift focus to any number of areas of interest like sports, music, technology or whatever 


* Examples of Open Data related to California Schools that we might work with initially include: 
    * [California Physical Fitness Test (PFT) program results:](https://www.cde.ca.gov/ta/tg/pf/pftresearch.asp)
    * [School metadata:](ttps://www.cde.ca.gov/ds/si/ds/pubschls.asp) 
    * [Academic performance data:](https://www.cde.ca.gov/ta/ac/cm/)
  
  
***

### Introductory Example

* The California Physical Fitness Test (PFT) program results data are a very interesting and important data set.  The data are broken down to the school, district and county levels, and include results of multiple physical tests and "body mass index". Because of this, the data set can be used to explore many types of questions and to provide insightful and valuable analyses and visual displays.

* Part of the R code for working with these data, and some R code and resultant charts, is shown below.  

```{r, echo=FALSE,include=FALSE}
library(dplyr)
library(readr)


fitnessgram_data   <- read_csv("Fitnessgram_data/2017_18_ResearchFile.txt",na = c("", "**")) # use as needed: ,col_types="_ciiiii") 
fitnessgram_labels <- read_csv("Fitnessgram_data/2017_18_Entities.txt")


fg_districts_data  <- filter(fitnessgram_data,
                            Level_Number==2,        # District
                            Report_Number==0)  %>%  # All Students   
                      mutate(link=paste0(DIST,ChrtNum))
                                       
district_codes_names <- filter(fitnessgram_labels, scode=="0000000", dcode !="00000") %>%
                          mutate(link=paste0(dcode,chrtnum)) %>%
                          select(link,chrtnum, dcode, District, County)

fg_districts_info     <- left_join(fg_districts_data,
                         district_codes_names,
                         by= "link")

```





```{r, warning=FALSE}
test_selected    <- "Upper Body Strength"
county_selected  <- "Contra Costa"                         
selected_info    <- filter(fg_districts_info, 
                           Line_Text == test_selected, 
                           County == county_selected, 
                           NoStud5 != 0)

hist(selected_info$Perc5a, breaks=20,
     main=paste("Histogram of",test_selected,"in",county_selected,"Schools"),cex.main=0.9,
     xlab = "Score"
)

```



* This chart above might lead to other questions, like, what about the other measures of fitness? What about my specific school? What about my school district?  What about boys compared to girls?  Can I see this on a map?  How can we know if the differences we see are "real" or just do to chance?  We could address all these questions together with the data we already have. A partial answer to the first question is below.  




```{r, warning=FALSE}
library(ggplot2)
county_selected  <- "Contra Costa"                         
selected_info    <- filter(fg_districts_info, 
                           County == county_selected, 
                           Table_Number == 1,
                           NoStud5 != 0)


ggplot(selected_info, aes(x=Perc5a)) + 
        geom_histogram(binwidth=4,colour="white") +
        facet_wrap( ~ Line_Text, ncol=2)    
```


### Beginning of Resource List

* R markdown resources:

    * https://stackoverflow.com/questions/29787850/how-do-i-add-a-url-to-r-markdown/29788494

