---
title: "Map Example"
author: "Michael Samuel"
date: "January 22, 2019"
output:
  html_document: default
  word_document: default
---

*  Michael Samuel
*  samfam921@gamil.com
*  Michael.Samuel@cdoph.ca.gov
*  https://mcsamueldatasci.github.io/Data_Science_Team/
*  https://github.com/mcSamuelDataSci/Data_Science_Team


<br>

The library() function loads "packages" that perfrom special functions into the workspace.

*  the dplyr packages is a workhorse for data manipulation
*  the tigris package connect to the US census and reads in geographic "shape" files.  It requires a free "census key" to run
*  tmap is the R package I use for mapping. There are many others.
<br>

```{r message=FALSE}
library(dplyr)   
library(tigris)  
library(tmap)
options(tigris_class = "sf")  # Read shape files as Simiple Features objects
```

This step reads in a dataset from the cloud on votting patterns in Califonira. The dataset is from a California Open Data Set.  The main portal home pages is here https://data.chhs.ca.gov/

```{r}
voterData.t <- read.csv("https://data.chhs.ca.gov/dataset/40fd0792-2bfd-4303-a848-fc5cb4338295/resource/c384c86a-49d2-4128-8389-b2701ff0bc35/download/voter-registration-2002-2010.csv",as.is=TRUE)

```

This little step reads in a census-tract level geographic "shape file" for California from the US census. (A free "key" from the US Census in needed if you want to run it yourself; available here https://api.census.gov/data/key_signup.html. )

```{r results='hide'}
tracts_CA   <- tracts(state = "CA", cb = TRUE) 
```


```{r}
voterData <- filter(voterData.t, geotype=="CT", 
                                 reportyear==2010, 
                                 race_eth_name=="Total", 
                                 type == "voted/registered") %>%
             transform(GEOID=paste0("0",geotypevalue))

```

This step merges the data using "join"

```{r warning=FALSE}
map.1    <- left_join(tracts_CA, voterData, by="GEOID")
```

This step makes a very simple map, with coloring based on the percent in the census tract who voted

```{r  warning=FALSE}
tmap_mode("view")
tm_shape(map.1) + tm_polygons(col="percent")
```

This step filter the data to just Conta Costa County, and re-maps

```{r warning=FALSE }
map.2 <- filter(map.1,county_name == "Contra Costa")

tm_shape(map.2) + tm_polygons(col="percent")
```











